The {\sc VSPEC} code: Spectroscopic phase curves of 3D exoplanet models in the presence of stellar variability

### 2024-06-04 src/tex/ms.tex
Copy up to examples section

>>>>

\subsection{Flares \label{subsec:flares}}
Flares are an important source of stellar variability on short timescales. We use the {\sc xoflares}
package \citep{barclay2020} to generate realistic lightcurve shapes and assume a blackbody spectrum.
{\sc xoflares} uses a flare profile function empirically derived by \citet{davenport2016} to
describe a flare lightcurve. Temperature, $E$, FWHM, and the peak time together completely describe a flare's multiwavelength lightcurve.

Before simulating an observation, {\sc VSPEC} pre-computes a population of flares based on the frequency-energy relationship from \citet{gao2022}:
\begin{equation} \label{eq:flare_freq}
    \log{(f{\rm [day]})} = \beta + \alpha~\log{(E/{\rm [erg]})}
\end{equation}

Where $f$ is the frequency of flares with energies $\ge E$.
We compute the number of expected flares over a time duration and determine the number to create $N$ by a random Poisson draw.
We then generate $N$ flares with energies determined by the quantile function:
\begin{equation}
    E = E_{\text{min}} (1-X)^{1/\alpha}
\end{equation}

where $E_{\text{min}}$ is the minimum considered flare energy and $X$ is a random number on the interval $[0,1)$.
We set the default values $\alpha=-0.829$, $\beta=26.87$ from \citet{gao2022}. Figure \ref{fig:flare_freq}
shows our simulation results compared to their powerlaw.

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{figures/flare_freq.pdf}
    \caption{
        {\bf Top}: Generated flare frequencies compared to those expected from \citet{gao2022}, generated using a 10,000 day simulation.
        The $y$-axis shows the frequency of flares with energies {\em greater than or equal to} the value of the $x$-axis. Error bars
        based on the square root of the number of observed flares as this is a Poisson process.
        {\bf Bottom}: Lightcurve of a star flaring with the same powerlaw slope as the top panel, but the intercept ($\beta$) has been increased by 0.3
        for visual effect. The mean flare temperature is $9000$ K and the mean FWHM is $3$ hours.
        }
    \script{flare_freq.py}
    \label{fig:flare_freq}
\end{figure}

\subsection{Granulation}
Granulation is a source of stellar variability that arises from convection near the surface of the star.
The result is a stellar surface that is not constant in temperature and that changes on very short timescales.
Hydrodynamic simulations of stellar atmospheres \citep[e.g.][]{magic2014} describe hot granules of rising gas surrounded by cooler,
sinking regions with a \teff a few percent lower than the nominal value.

We model granulation as a global process, and its effects are computed after the effects of spots and faculae.
Of the remaining ``quiet'' photosphere (i.e. the regions not covered in spots of faculae), a fraction is computed to be part of the cool
inter-granule surface. The surface coverage of the cool region at any given time is computed by a Gaussian process (GP) using
the \sc{tinygp} package \citep{foreman-mackey2024} following the methodology of \citet{gordon2020}.
The GP uses a custom kernel function based on a power spectrum \citep{anderson1990,kallinger2014} to produce random changes in the coverage.

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{figures/granulation_lightcurve.pdf}
    \caption{
        Effect of granulation on the lightcurve of a star. This simulation assumes a mean
        10\% coverage by inter-granule region whose temperature is 300 K lower than the surrounding photosphere.
        It also assumes that the coverage value varies with a magnitude of 1\% on a timescale of 6 hours.
        }
    \script{granule_lc.py}
    \label{fig:gran_lc}
a\end{figure}

\section{Planet Model \label{sec:pl_model}}
Our model uses the Planetary Spectrum Generator\footnote{\href{https://psg.gsfc.nasa.gov/}{https://psg.gsfc.nasa.gov/}}
\citep[PSG,][]{villanueva2018} to calculate the planetary spectrum. We upload a 3D Global Circulation Model (GCM) to PSG's
Global Emission Spectra (GlobES) application to retrieve thermal and reflected spectra for various points in the planet's phase curve. \TJ{cite Vincent's paper}

\subsection{Interfacing through {\sc pypsg}}
\ref{subsec:pypsg}
{\sc VSPEC} interfaces with PSG using the {\sc pypsg} package \TJ{link}, which is essentially a Python wrapper for the PSG API system.

\TJ{Write much more}

\subsection{3D Planet models}
{\sc VSPEC} allows the user to upload any desired GCM. However, for questions that focus on detectability,
the specific physics of the GCM are less important than properties like day/night temperature, surface pressure, or
the major absorbers in the atmosphere. To simplify these studies, {\sc VSPEC} has a built-in GCM that takes in just a
few parameters and produces a simple-yet-robust exoplanet climate. Simplicity allows the user to know everything about
the model and reduce surprises, but its basis in thermodynamics and robust energy balance make this model fit for publication-level science \TJ{is this really true? probably not}.

The {\sc VSPEC} GCM centers around a 2D temperature map of the planet's surface. Based on \citet{cowan2011}, this map
is constructed completely given the incident flux from the star (parameterized by the host \teff and radius
and the planet's semimajor axis and Bond albedo) and the unitless thermal redistribution efficiency $\epsilon$.
This efficiency is 0 for a planet that reradiates instantly and $\gg 1$ for an atmosphere that efficiently redistributes heat to its night side.
Figure \ref{fig:thermal_inertia} shows a temperature map example for an Earth-like planet with $\epsilon = 2\pi$.
These temperature maps are validated by the energy balance of the planet; {\sc VSPEC} computes the incident and emergent
flux across the planet and asserts that they agree.

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{figures/cowan_gcm.pdf}
    \caption{
        Example of a surface temperature map created by {\sc VSPEC}. The solar flux at Earth,
        a Bond albedo of 0.3, and $\epsilon = 2\pi$ were used in this case.
        Notice that the hottest point is offset from the sub-stellar point due to thermal inertia.
    }
    \script{cowan_gcm.py}
    \label{fig:thermal_inertia}
\end{figure}

Now that we have the surface temperature at each GCM coordinate we can build the atmosphere on top of it.
We choose a surface pressure for our model and generate a pressure profile as a function of layer. It is
important to note that the choice of the pressure profile -- provided it is scaled to the surface pressure and sampled in a reasonable manner -- is
arbitrary and will not affect the spectrum of the planet. This is because PSG calculates the altitude and thickness of each layer based on the
pressure and surface gravity; the pressure profile is essentially a proxy for the sampling of atmosphere layers as a function of altitude.

Next we create an adiabatic temperature profile at each GCM coordinate. An adiabat assumes that
\begin{equation}
    T^\gamma P^{1 - \gamma} = \text{const.}
\end{equation}
where $\gamma$ is the adiabatic index. Therefore the temperature profile can be computed
\begin{equation}
    T = T_{\rm surf} \left(\frac{P}{P_{\rm surf}} \right)^{1 - \frac{1}{\gamma} }
\end{equation}

Finally, we populate our atmosphere with molecules. The user specifies the species and abundance and atmosphere is filled accordingly with a constant molecular suite.

The caveats of this approach lie in the simplified physics of the atmosphere structure. The heat-redistribution model does not
capture the effects of latitudinal mixing; an adiabatic temperature profile will never exhibit the temperature inversion
that we know form in complex atmospheres like Earth's. Similarly, setting a constant abundance does not allow for water vapor to
vary with temperature as it does on Earth. We must also be careful not to add so much of an absorbing species that the energy balance of
the atmosphere would change; instead we treat them as trace species in an otherwise transparent atmosphere. Our simplified approach
allows us to easily examine a large parameter space of atmospheres whose behaviors are easy to predict.
